---
title: "Trying Something Else"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)

```
Open and rename Data
```{r}
prgusap1_puf <- read_sas("prgusap1_puf.sas7bdat")
View(prgusap1_puf)
df <- prgusap1_puf
```
Cut out Extra variables and rename columns
```{r}
#select(RACETHN_5CAT, REGION_US, EDCAT8, A_N01_T)
#rename(Race5=RACETHN_5CAT, Region=REGION_US, EDU8=EDCAT8, Gender= A_N01_T)
df_smallg <- df%>%
  select(RACETHN_5CAT, REGION_US, EDCAT8, A_N01_T) %>%
  rename(Race=RACETHN_5CAT, Region=REGION_US, EDUC=EDCAT8, Gender=A_N01_T)

```
Change name of Race Variables
```{r}
df_Raceg <- df_smallg %>%
  mutate(Race = ifelse(Race==1, "Hispanic",
                       ifelse(Race==2, "White",
                              ifelse(Race==3, "Black",
                                     ifelse(Race==4, "Asian/Pacific Islander",
                                            ifelse(Race==6, "Other", "Not Stated" ))))))
                                                   
```
Change name of Region Variables
```{r}
df_Region_Raceg <- df_Raceg %>%
  mutate(Region = ifelse(Region== 1, "Northeast",
                         ifelse(Region== 2, "Midwest",
                                ifelse(Region== 3, "South",
                                       ifelse(Region== 4, "West", "Other Region")))))
```
Change name of Education Variables
```{r}
df_Ed_Reg_Raceg <- df_Region_Raceg %>%
  mutate(EDUC = ifelse(EDUC==1 | EDUC==2, "Primary",
                       ifelse(EDUC==3 | EDUC==4, "Secondary",
                              ifelse(EDUC==6, "Bachelors",
                                     ifelse(EDUC==7, "Masters",
                                            ifelse(EDUC==5 | EDUC==8, "Professional", "Other"))))))
                                                   
```
Change name of Gender Variables
```{r}
df_4VAR <- df_Ed_Reg_Raceg %>%
  mutate(Gender = ifelse(Gender== 1, "Male",
                         ifelse(Gender== 2, "Female",
                                 "Other ")))
```
Descriptive Statistics Counts and Percent All
```{r}
desc_stat_ALLg <- df_4VAR %>%
  group_by(Region, Race, EDUC, Gender)%>%
  summarize(N=n())%>%
  ungroup()%>%
  group_by(Region) %>%
  mutate (perc = N/sum(N))
  
write.csv(desc_stat_ALLg, "desc_stat_ALLg.csv")
desc_stat_ALLg <- as.data.frame(desc_stat_ALLg)
```
Descriptive Statistics for plot for Gender by Region
```{r}
desc_stat_Genderg <- df_4VAR %>% group_by(Gender,Region) %>%
  summarise(N=n()) %>%
  ungroup() %>%
  group_by(Region) %>%
  mutate (perc = N/sum(N))

write.csv(desc_stat_Genderg, "desc_stat_Genderg.csv")
desc_stat_Genderg <- as.data.frame(desc_stat_Genderg)
```
Plot of Gender by Region
```{r}
ggplot(desc_stat_Genderg, aes(x = Region, y=perc, fill=Gender)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(label = scales::: percent_format(accuracy =1))
```
Descriptive Statistics for Plot of Race by Region
```{r}
desc_stat_Raceg <- df_4VAR %>% group_by(Region,Race) %>%
  summarise(N=n()) %>%
  ungroup() %>%
  group_by(Region) %>%
  mutate (perc = N/sum(N))

write.csv(desc_stat_Raceg, "desc_stat_Raceg.csv")
desc_stat_Raceg <- as.data.frame(desc_stat_Raceg)
```

Plot of Race by Region
```{r}
ggplot(desc_stat_Raceg, aes(x = Region, y=perc, fill=Race)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(label = scales::: percent_format(accuracy =1))
```
Descriptive Statistics of Education by Region
```{r}
desc_stat_EDUCg <- df_4VAR %>% group_by(EDUC, Region) %>%
  summarise(N=n()) %>%
  ungroup() %>%
  group_by(Region) %>%
  mutate (perc = N/sum(N))

write.csv(desc_stat_EDUCg, "desc_stat_EDUCg.csv")
desc_stat_EDUCg <- as.data.frame(desc_stat_EDUCg)
```

Plot of Education by Region
```{r}
desc_stat_EDUCg$EDUC <- factor(desc_stat_EDUCg$EDUC, levels =c("Primary", "Secondary", "Bachelors", "Masters", "Professional", "Other"))

ggplot(desc_stat_EDUC, aes(x = Region, y=perc, fill=EDUC)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(label = scales::: percent_format(accuracy =1))
```
Take out NA in all but Region
```{r}
df_cleang <- df_4VAR %>%
  filter(!row_number() %in% c(29, 45, 54, 102, 112, 131, 139, 140, 196, 218, 231, 248, 249, 266, 380, 398, 412, 419, 440, 449, 533, 542, 582, 586, 590, 597, 615, 624, 636, 660, 675, 684, 711, 747, 755, 766, 820, 825, 832, 839, 861, 891, 897, 940, 942, 951, 959, 964, 1003, 1046, 1053, 1057, 1077, 1080, 1097, 1118, 1119, 1129, 1142, 1155, 1270, 1284, 1287, 1288, 1303, 1345, 1355, 1367, 1404, 1408, 1422, 1438, 1474, 1484, 1577, 1593, 1622, 1634, 1637, 1640, 1664, 1665, 1676, 1710, 1712, 1739, 1751, 1767, 1813, 1862, 1864, 1900, 1931, 1942, 1967, 1975, 1996, 1999, 2007, 2027, 2043, 2058, 2070, 2071, 2083, 2086, 2093, 2101, 2102, 2123, 2133, 2139, 2183, 2184, 2212, 2216, 2238, 2242, 2258, 2266, 2278, 2318, 2358, 2372, 2374, 2384, 2479, 2483, 2487, 2490, 2493, 2546, 2550, 2554, 2558, 2587, 2598, 2607, 2618, 2656, 2689, 2694, 2737, 2755, 2771, 2790, 2816, 2839, 2878, 2880, 2919, 2925, 2935, 2952, 2966, 2977, 2979, 3011, 3114, 3144, 3161, 3164, 3176, 3224, 3227, 3235, 3278, 3284, 3493, 3497, 3563, 3568, 3576, 3602, 3615, 3627, 3634, 3644, 3653, 3655))

write.csv(df_cleang, "df_cleang.csv")
```
Descriptive Statistics for plot for Gender by Region w/o 3 NA obs
```{r}
desc_stat_Gender_cleang <- df_cleang %>% group_by(Region,Gender) %>%
  summarise(N=n()) %>%
  ungroup() %>%
  group_by(Region) %>%
  mutate (perc = N/sum(N))

write.csv(desc_stat_Gender_cleang, "desc_stat_Gender_cleang.csv")
desc_stat_Gender_cleang <- as.data.frame(desc_stat_Gender_cleang)
```
Plot of Gender by region w/o 3 NA obs
```{r}
ggplot(desc_stat_Gender_cleang, aes(x = Region, y=perc, fill=Gender)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(label = scales::: percent_format(accuracy =1))

```
Descriptive Statistics for plot of Race by Region w/o 3NA obs
```{r}
desc_stat_Race_cleang <- df_cleang %>% group_by(Region,Race) %>%
  summarise(N=n()) %>%
  ungroup() %>%
  group_by(Region) %>%
  mutate (perc = N/sum(N))

write.csv(desc_stat_Race_cleang, "desc_stat_Race_cleang.csv")
desc_stat_Race_cleang <- as.data.frame(desc_stat_Race_cleang)
```
Plot of Race by Region w/o 3 NA obs
```{r}
ggplot(desc_stat_Race_cleang, aes(x = Region, y=perc, fill=Race)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(label = scales::: percent_format(accuracy =1))
```
Descriptive Statistics of Education by Region w/o 3 NA obs
```{r}
desc_stat_EDUC_cleang <- df_cleang %>% group_by(EDUC, Region) %>%
  summarise(N=n()) %>%
  ungroup() %>%
  group_by(Region) %>%
  mutate (perc = N/sum(N))

write.csv(desc_stat_EDUC_cleang, "desc_stat_EDUC_cleang.csv")
desc_stat_EDUC_cleang <- as.data.frame(desc_stat_EDUC_cleang)
```
Plot of Education by Region
```{r}
desc_stat_EDUC_cleang$EDUC <- factor(desc_stat_EDUC_cleang$EDUC, levels =c("Primary", "Secondary", "Bachelors", "Masters", "Professional", "Other"))

ggplot(desc_stat_EDUC_cleang, aes(x = Region, y=perc, fill=EDUC)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(label = scales::: percent_format(accuracy =1))
```
Load Packages We Need
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mice)

```
Fix variable type (as.factor())
```{r}
df_asfactorg <- df_cleang %>%
  mutate(Region = as.factor(Region),
         Race = as.factor(Race),
         EDUC = as.factor(EDUC),
         Gender = as.factor(Gender))
summary(df_asfactorg)
```
Trying mice
```{r}
df_impg <- mice(df_asfactorg, m = 2, seed = 1)
```
remove problems
```{r}
pred <- quickpred(df_asfactorg)
pred[,] <- 0
df_impg <- mice(df_asfactorg, m= 2, seed= 1, predictorMatrix = pred)
```
check
```{r}
summary(df_impg)

complete(df_impg)

summary(complete(df_impg))
```
Save this file
```{r}
save(df_impg, file= "df_impg_clean.rda")
```