---
title: "Trying Something Else"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(mice)
library(lme4)
library(MuMIn)
library(mitml)
library(miceadds)
library(AICcmodavg)
```
Open and rename Data
```{r}
prgusap1_puf <- read_sas("prgusap1_puf.sas7bdat")
View(prgusap1_puf)
df <- prgusap1_puf
```
Cut out Extra variables and rename columns
```{r}
#select(RACETHN_5CAT, REGION_US, EDCAT8, A_N01_T)
#rename(Race5=RACETHN_5CAT, Region=REGION_US, EDU8=EDCAT8, Gender= A_N01_T)
df_smallg <- df%>%
  select(RACETHN_5CAT, REGION_US, EDCAT8, A_N01_T, J_Q09USX) %>%
  rename(Race=RACETHN_5CAT, Region=REGION_US, EDUC=EDCAT8, Gender=A_N01_T, Income=J_Q09USX)

```
Change name of Race Variables
```{r}
df_Raceg <- df_smallg %>%
  mutate(Race = ifelse(Race==1, "Hispanic",
                       ifelse(Race==2, "White",
                              ifelse(Race==3, "Black",
                                     ifelse(Race==4, "Asian/Pacific Islander",
                                            ifelse(Race==6, "Other", "Not Stated" ))))))
                                                   
```
Change name of Region Variables
```{r}
df_Region_Raceg <- df_Raceg %>%
  mutate(Region = ifelse(Region== 1, "Northeast",
                         ifelse(Region== 2, "Midwest",
                                ifelse(Region== 3, "South",
                                       ifelse(Region== 4, "West", "Other Region")))))
```
Change name of Education Variables
```{r}
df_Ed_Reg_Raceg <- df_Region_Raceg %>%
  mutate(EDUC = ifelse(EDUC==1 | EDUC==2, "Primary",
                       ifelse(EDUC==3, "Secondary",
                              ifelse(EDUC ==4 | EDUC==5, "Professional",
                              ifelse(EDUC==6, "Bachelors",
                                     ifelse(EDUC==7, "Masters",
                                             ifelse(EDUC==8, "Research", "Other")))))))
                                                   
```
Change name of Gender Variables
```{r}
df_4VAR <- df_Ed_Reg_Raceg %>%
  mutate(Gender = ifelse(Gender== 1, "Male",
                         ifelse(Gender== 2, "Female",
                                 "Other ")))
```
Income
```{r}
df_5VAR <- df_4VAR %>%
  mutate(Income_cont = ifelse(Income== 1, "5000",
                         ifelse(Income== 2, "10000",
                         ifelse(Income== 3, "25000", 
                         ifelse(Income== 4, "35000",
                         ifelse(Income== 5, "45000",
                          ifelse(Income== 6, "55000",
                          ifelse(Income== 7, "67500",
                          ifelse(Income== 8, "87500",
                          ifelse(Income== 9, "125000",
                          ifelse(Income== 10, "200000", 
                          ifelse(Income== 11, "0", "Other")))))))))))) 
```
Mutate
```{r}
df <- df_5VAR %>%
  mutate(Income = ifelse(Income== 1, "1-9,999",
                  ifelse(Income== 2, "10,000-19,999",
                  ifelse(Income== 3, "20,000- 29,999", 
                  ifelse(Income== 4, "30,000-39,999",
                   ifelse(Income== 5, "40,000- 49,999", 
                  ifelse(Income== 6, "40,000-59,999",
                   ifelse(Income== 7, "60,000-74,999",
                   ifelse(Income== 8, "75,000-99,999",
                    ifelse(Income== 9, "100,000-149,999",
                    ifelse(Income== 10, "150,000 +", 
                    ifelse(Income== 11, "0", "Other" ))))))))))))
         
```

```{r}
df <- df %>%
  mutate(Region_hlm = ifelse(Region == "Northeast", 1,
                           ifelse(Region== "Midwest", 2,
                                  ifelse(Region == "South", 3,
                                         ifelse(Region== "West", 4, "Other")))))
```

Identify where there are missing variables & Fix variable type (as.factor())
```{r}

        
df$Income_cont <- as.numeric(df$Income_cont)
df$Region_hlm <- as.numeric(df$Region_hlm)
df$Region <- as.factor(df$Region)
df$Income <- as.factor(df$Income)
df$EDUC <- as.factor(df$EDUC)
df$Gender <- as.factor(df$Gender)
df$Race <- as.factor(df$Race)

summary(df)

```

Trying mice M=# of times imputed
```{r}
df_imp <- mice(df, m = 10, seed = 1)
```
remove problems
```{r}
pred <- quickpred(df_asfactorg)
pred[,] <- 0
df_impg <- mice(df_asfactorg, m= 2, seed= 1, predictorMatrix = pred)
```
check
```{r}
summary(df_imp)

complete(df_imp)

summary(complete(df_imp))
```
Save this file
```{r}
save(df_imp, file= "df_imp.rda")
```
Load the data, make complete dataset, and mitml
```{r}
load("~/ResEv Class Project/df_imp.rda")
df <- complete(df_imp)
df_mitml <- mids2mitml.list(df_imp)
save(df, file= "df.rda")
```
```{r}
library(mice)
library(lme4)
library(AICcmodavg)
```
```{r}
#load your data
load("~/ResEv Class Project/df.rda")
#complete for MI
df <- complete(df)
```
```{r}
#run models
mod1 <- with(df_mitml, lm(Income_cont ~ 1 + Race + Region + Gender + EDUC, df))
mod2 <- with(df_mitml, lm(Income_cont ~ 1 + Race*Gender + Region + EDUC, df))
mod3 <- with(df_mitml, lm(Income_cont ~ 1 + Race*Gender*Region + EDUC, df))
mod4 <- with(df_mitml, lm(Income_cont ~ 1 + Race*Region + Gender + EDUC, df))
mod5 <- with(df_mitml, lm(Income_cont ~ 1 + Gender*Region + Race + EDUC, df))
mod6 <- with(df_mitml, lm(Income_cont ~ 1 + Region*EDUC + Race + Gender, df))
mod7 <- with(df_mitml, lm(Income_cont ~ 1 + Region*EDUC + Race*Gender, df))
mod8 <- with(df_mitml, lm(Income_cont ~ 1 + Region*EDUC*Race*Gender, df))

summary(mod1)
```
```{r}
#See AIC scores
#define list of models
models <- list(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8)
#specify model names
mod.names <- c("mod1","mod2","mod3","mod4","mod5","mod6","mod7","mod8")

#calculate AIC of each model
aictab(cand.set = models, modnames = mod.names)

```
Run models
```{r}
#mod_1 <- lm(EDUC_cont ~ 1, df)
#mod_2 <- lm(EDUC_cont ~ 1 + Race, df)
#mod_3 <- lm(EDUC_cont ~ 1 + Race + Region, df)
#mod_4 <- lm(EDUC_cont ~ 1 + Race*Region, df)
  

mod1 <- with(df_mitml, lm(Income_cont ~ 1 + Race + Region + Gender + EDUC, df))
mod2 <- with(df_mitml, lm(Income_cont ~ 1 + Race*Gender + Region + EDUC, df))
mod3 <- with(df_mitml, lm(Income_cont ~ 1 + Race*Gender*Region + EDUC, df))
mod4 <- with(df_mitml, lm(Income_cont ~ 1 + Race*Region + Gender + EDUC, df))
mod5 <- with(df_mitml, lm(Income_cont ~ 1 + Gender*Region + Race + EDUC, df))
mod6 <- with(df_mitml, lm(Income_cont ~ 1 + Region*EDUC + Race + Gender, df))
mod7 <- with(df_mitml, lm(Income_cont ~ 1 + Region*EDUC + Race*Gender, df))
mod8 <- with(df_mitml, lm(Income_cont ~ 1 + Region*EDUC*Race*Gender, df))
```
Check output
```{r}

testEstimates(mod1, extra.pars=TRUE)
testEstimates(mod2, extra.pars=TRUE)
testEstimates(mod3, extra.pars=TRUE)
testEstimates(mod4, extra.pars=TRUE)
testEstimates(mod5, extra.pars=TRUE)
testEstimates(mod6, extra.pars=TRUE)
testEstimates(mod7, extra.pars=TRUE)
testEstimates(mod8, extra.pars=TRUE)
testEstimates(mod9, extra.pars=TRUE)
testEstimates(mod10, extra.pars=TRUE)
```