---
title: "Math imputation"
output: html_document
---

Load packages we need
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mice)
library(tidyverse)
```

Load data & make new df of it
```{r}
cleaned_data <- readRDS("~/RESEV553/Math analysis/cleaned_data.rds")

```

identify where we have missing data & character variables
```{r}

```

Fix variable type (as.factor())
```{r}




```

df <- df %>%
  mutate(S_Teacher = as.factor(S_Teacher),
         F_Teacher = as.factor(F_Teacher),
         School = as.factor(School),
         State = as.factor(State),
         Race = as.factor(Race),
         ESL = as.factor(ESL),
         F_Per = as.factor(F_Per),
         S_Per = as.factor(S_Per),
         Gender = as.factor(Gender),
         IEP = as.factor(IEP),
         Lunch = as.factor(Lunch))

Simplest thing is to just let mice do it's thing
```{r}



```
Remove problems
```{r}

```



Check what happened
```{r}

```

Fix up any data afterwards
```{r}


# MIdata<-mids2mitml.list(df_imp) #converts file type
# thing <- list()
# for (i in 1:10){
#   temp <- MIdata[[i]]
#   temp <- temp %>%
#     merge(cleaned_data$ID)
#   thing[[i]] <- temp
#   }
# MIdata <- as.mitml.list(thing)
```


Save the file
```{r}

```



Load packages we need
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mice)
library(tidyverse)
```

Load data & make new df of it
```{r}
cleaned_data <- readRDS("~/RESEV553/Math analysis/cleaned_data.rds")


#df <- cleaned_data
```

identify where we have missing data & character variables
```{r}
summary(df)
```

Create indicator variables (will need this for future stuff)
```{r}

df <- df %>%
  mutate(S_Teacher = as.factor(S_Teacher),
         F_Teacher = as.factor(F_Teacher),
         School = as.factor(School),
         State = as.factor(State),
         Race = as.factor(Race),
         ESL = as.factor(ESL),
         F_Per = as.factor(F_Per),
         S_Per = as.factor(S_Per),
         Gender = as.factor(Gender),
         IEP = as.factor(IEP),
         Lunch = as.factor(Lunch),
        na_count = apply(is.na(df), 1, sum))# %>%
  # select(-ID)


```



Simplest thing is to just let mice do it's thing
```{r}
df_imp <- mice(df, m=2, seed = 1)
```
Remove problems
```{r}
pred <- quickpred(df)
pred[,"ID"] <- 0
df_imp <- mice(df, m=2, seed = 1, predictorMatrix = pred)
```



Check what happened
```{r}
summary(df_imp)
complete(df_imp)
summary(complete(df_imp))
```

Fix up any data afterwards
```{r}
MIdata<-mids2mitml.list(df_imp) #converts file type
thing <- list()
for (i in 1:10){
  temp <- MIdata[[i]]
  temp <- temp %>%
    merge(cleaned_data$ID)
  thing[[i]] <- temp
  }
MIdata <- as.mitml.list(thing)
```


Save the file
```{r}
save(df_imp, file="df_imp.rds")
```